Multidimensional signals and analytic flexibility: Estimating degrees of freedom in human speech analyses

### 2022-09-09 RR_manuscript/Draft_RR.Rmd
♻️ - edit descriptive results

# acoustic operationalizations
# Gross operationalization of acoustic operationalization
unique_acou_oper <- msa_models %>%
  unite("unique_oper", outcome, operationalisation, operationalisation_notes, na.rm = TRUE) %>%
  count(unique_oper)

@@

@@

In the following, we will describe the characteristics of our analyst teams and their models in order to give an overview over the sample and the large variation across analysis strategies.

@@

### Team properties

`r team_desc_list$Initial$n_teams$val %>% xfun::numbers_to_words() %>% stringr::str_to_title()` teams initially signed up to participate. 
`r (team_desc_list$Initial$n_teams$val - team_desc_list$Final$n_teams$val) %>% xfun::numbers_to_words() %>% stringr::str_to_title()` of the proposed teams dropped out of the project during the analysis phase^[ADD REASONS]. 
The remaining `r team_desc_list$Final$n_teams$val` teams submitted their analysis in time and were coded by the three initiating authors (SC, JC, TR).

Submitting teams consisted of an average of `r team_desc_list$Final$mean_size$val` analysts (sd = `r team_desc_list$Final$sd_size$val`). 
Analysts had on average `r round(phd_mean[[1]],1)` years (sd = `r round(phd_mean[[2]],1)`) post PhD experience, ranging from `r round(phd_mean[[3]],1)` years, i.e. PhD students (or younger) to `r round(phd_mean[[4]],1)` years.

@@

Analysts prior belief in the effect under investigation ranged from `r round(prior_mean[[3]],1)`  to `r round(prior_mean[[4]],1)` with an average of `r round(prior_mean[[1]],1)` (sd = `r round(prior_mean[[2]],1)`), indicating that analysts had an overall rather high prior plausibility of the investigated relationship between prosody and typicality (scale ranged from 0 to 100).

The teams have submitted `r no_analyses` models to answer the research question. 
On average, teams submitted `r round(mean_analyses[[1]], 1)` models. 
Crucially, analytic approaches widely varied in how they acoustically analysed the speech signal and how the statistically analysed the extracted values.

@@

Teams differed in what and how they measured the acoustic signal, including choosing different aspects of the acoustic signal, the temporal window over which they measured, and the concrete operationalization of how they measured those acoustic properties.
`r round(outcome_n[[1,1]],2) * 100`\% of models used a duration measure as the outcome variable, `r round(outcome_n[[2,1]],2) * 100`\% used a f0 measure, `r round(outcome_n[[3,1]],2) * 100`\% used a formant measure, `r round(outcome_n[[4,1]],2) * 100`\% used an intensity measure, and `r round(outcome_n[[5,1]],2) * 100`\% used a different measure. <!-- TR: Do we want to list some of them?-->

`r round(window_n[[1,1]],2) * 100`\% of models measured acoustic properties at the level of the segment (e.g. comparing the acoustic profile of a vowel), `r round(window_n[[2,1]],2) * 100`\% at the level of the word (e.g. comparing the acoustic profile of "banana"), `r round(window_n[[3,1]],2) * 100`\% at the level of the phrase (e.g. the noun phrase including determiner and adjective, e.g. "the green banana"), `r round(window_n[[4,1]],2) * 100`\% at the level of the whole sentence, and `r round(window_n[[5,1]],2) * 100`\% used a different time window. Based on a coarse coding of how acoustic measures were operationalized, we find a total of `r nrow(unique_acou_oper)` different measurement specifications.

<!-- SC: I have added the number of unique meas. specs but we might want to discuss this. -->
<!-- TR: How did you do this?-->

@@

The multiverse of acoustic measurement choices is exponentiated with choices during the statistical analysis, including the predictors and their operationalization, the chosen inferential framework, the type of model, and their model architecture, specifically how they accounted for dependencies in the data via random effect specifications in their multilevel models.

On average, models included `r round(predictor_n,1)` different predictors (counting only conceptually different predictors), i.e. in addition to the critical predictor, typicality of the adjective noun combinations, most teams included additional predictors in their models. Predictors included the information structure of the sentence, trial number, semantic dimensions of the referent, part of speech, or gender of the speaker.

The original dataset allowed to operationalize the most relevant predictor, typicality, in different ways. 
It was operationalized as categorical (e.g. typical vs. atypical) in `r round(typicality_n[[1,1]],2) * 100`\% of models, continuous (on a scale from 0-100) with the mean typicality rating in `r round(typicality_n[[2,1]],2) * 100`\% of models, and continuous with the median typicality rating in `r round(typicality_n[[3,1]],2) * 100`\% of models.

The majority of models were rooted in the frequentist framework (`r round(framework_n[[2,1]],2) * 100`\%). `r round(framework_n[[1,1]],2) * 100`\% were operating in a Bayesian framework. 
While teams almost exclusively used linear models to analyze their data (`r round(model_n[[2,1]],2) * 100`\%, with a few exception using machine learning techniques or GAMs as a special case of a linear model), teams differed drastically in how they accounted for dependencies within the data. 
The data contains several dependencies between data points, with multiple data points coming from the same subject and with multiple data points being associated with the same adjective or noun. 
The appropriate way to account for this non-independence is by using multi-level models and specifying so-called random effects [e.g., @gelman2006data; @schielzeth2009conclusions]. 
`r round(random_n[[3,1]],2) * 100`\% of linear models specified no random effects at all (without pooling their data), so effectively ignoring these non-independences [@hurlbert1984pseudoreplication]. <!-- TODO: We need to check if these model pooled -->
`r round(random_n[[1,1]],2) * 100`\% specified random intercepts only, and `r round(random_n[[2,1]],2) * 100`\% specified random intercepts and random slopes to account for the non-independence.
On average, teams that specified random effects, included `r round(rterms_n,1)` random terms in their models. 
Based on statistical framework, type of model, distribution family, fixed terms, and discarding random effects, there were a total of `r nrow(unique_model_specs)` different model specifications. 
<!-- TR: If done on team level did every team have a different mdoel?  -->

@@

Teams reviewed each others' analyses for both the acoustic analysis and the statistical analysis. 
The mean rating of the quality of the acoustic analyses was `r round(review_mean[[1]],1)` (sd = `r round(review_mean[[2]],1)`); the mean rating of the quality of the statistical analysis was `r round(review_mean[[3]],1)` (sd = `r round(review_mean[[4]],1)`). For reference, the scale of the rating anchored a point grade of 75 as "an imperfect analysis but the needed changes are unlikely to dramatically alter final interpretation", indicating a rather high confidence of reviewers that the provided analyses yield appropriate (yet "imperfect") answers to the research question.

>>>>

@@

The results section is divided into three parts.
We first provide a statistical description of team composition, nature of acoustic analyses and statistical approaches, and peer-review ratings of these two as obtained by the team members.
Second, we report the results of the meta-analytic model, focussing on between-team variability and the estimated meta-analytic effect of word combination typicality on speech acoustics.
Finally, we present the analysis of the effect of analytic and researcher-related predictors on the meta-analytic effect.

@@

In the following sections, we will describe the characteristics of the analysis teams that participated in the study and the analytic approaches they adopted.
An important aspect that emerges from the descriptive analysis is the large variation in analytic strategies.

@@

### Characteristics of analytic teams

@@

`r team_desc_list$Initial$n_teams$val %>% xfun::numbers_to_words() %>% stringr::str_to_title()` teams initially signed up to participate in the study.
<!-- TODO: Add reasons for dropping out -->
`r (team_desc_list$Initial$n_teams$val - team_desc_list$Final$n_teams$val) %>% xfun::numbers_to_words() %>% stringr::str_to_title()` of the signed-up teams dropped out during the analysis phase^[ADD REASONS]. 
The remaining `r team_desc_list$Final$n_teams$val` teams submitted their analysis by the established deadline.
These are the teams that are included in this study.

Submitting teams consisted of an average of `r team_desc_list$Final$mean_size$val` individual analysts (SD = `r team_desc_list$Final$sd_size$val`).

Upon sign-up, we collected background information from each analyst through the intake form, which was administered during Phase 1, prior to the distribution of the data to the teams.
Analysts had on average `r round(phd_mean[[1]],1)` years (SD = `r round(phd_mean[[2]],1)`) of experience after completing their PhD, ranging from `r round(phd_mean[[3]],1)` years, i.e. PhD students (or younger) to `r round(phd_mean[[4]],1)` years.
The analysts' prior belief in the effect under investigation on a scale from 0 to 100 ranged from `r round(prior_mean[[3]],1)`  to `r round(prior_mean[[4]],1)` with an average of `r round(prior_mean[[1]],1)` (SD = `r round(prior_mean[[2]],1)`).
We take this to suggest that analysts had an overall rather high positive prior belief in the investigated relationship between prosody and word combination typicality.

At the end of Phase 2 (primary data analysis), the teams have submitted a grand total `r no_analyses` individual models to answer the research question, with a mean of `r round(mean_analyses[[1]], 1)` models per team. 
Crucially, the analytic approaches employed by the teams varied widely both in the acoustical analyses of the speech signal and in the statistical analyses of the obtained measures.
We proceed to discuss these in the following sections, after which we provide an overview of the peer-review ratings.

@@

The analytic teams differed in their approach to the acoustic analysis of the speech signal, including choosing specific acoustic measures, the temporal window over which they are taken from and how the measures are transformed.
`r round(outcome_n[[1,1]],2) * 100`\% of models used a measure of duration as the outcome variable, `r round(outcome_n[[2,1]],2) * 100`\% used fundamental frequency (f0), `r round(outcome_n[[3,1]],2) * 100`\% used vowel formants, `r round(outcome_n[[4,1]],2) * 100`\% intensity, and `r round(outcome_n[[5,1]],2) * 100`\% other measures.
<!-- TR: Do we want to list some of them?-->

```{r unique-acou-oper}
# acoustic operationalizations
# Gross operationalization of acoustic operationalization
unique_acou_oper <- msa_models %>%
  unite("unique_oper", outcome, operationalisation, operationalisation_notes, na.rm = TRUE) %>%
  count(unique_oper)
```

`r round(window_n[[1,1]],2) * 100`\% of models obtained acoustic measures at the level of the segment (e.g. comparing the acoustic profile of a vowel), `r round(window_n[[2,1]],2) * 100`\% from the word level (e.g. comparing the acoustic profile of *Banane* 'banana'), `r round(window_n[[3,1]],2) * 100`\% at the level of the phrase (e.g. the noun phrase including determiner and adjective, e.g. "the green banana"), `r round(window_n[[4,1]],2) * 100`\% from the whole sentence, and `r round(window_n[[5,1]],2) * 100`\% used a different time window.
Based on a coarse coding of how acoustic measures were operationalized, we find a total of `r nrow(unique_acou_oper)` different measurement specifications.

@@

The high variability of acoustic measurements is multiplied by the choices of statistical analysis, including the chosen inferential framework, the type of model, and the model specification, like choice of predictors, interactions and group-level (aka random) effects.

The models included a mean of `r round(predictor_n)` different predictors (defined as variables or columns in the data table).
This means that, in addition to the critical predictor (typicality of the adjective noun combinations), there was on average at least an additional predictor.
Among the predictors included in the models there were information structure of the sentence, trial number, semantic dimensions of the referent, part of speech, or speaker gender.

The data given to the teams allowed them to operationalize the predictor of interest, word typicality, in different ways.
Among the possible operationalizations, `r round(typicality_n[[1,1]],2) * 100`\% of models contained typicality as a categorical variable (e.g. atypical vs. typical), `r round(typicality_n[[2,1]],2) * 100`\%  used a continuous typicality scale from 0-100 by calculating the mean typicality for each word combination as obtained from the norming study, while `r round(typicality_n[[3,1]],2) * 100`\% of the models used the median typicality rating.

The majority of models were run within a frequentist framework (`r round(framework_n[[2,1]],2) * 100`\%).
`r round(framework_n[[1,1]],2) * 100`\% were run within a Bayesian framework.
While teams almost exclusively used linear models to analyze their data (`r round(model_n[[2,1]],2) * 100`\%, with a few exception using machine learning techniques or GAMs as a special case of linear models), teams differed drastically in how they accounted for dependencies within the data.

The data contains several dependencies between data points, with multiple data points coming from the same subject and with multiple data points being associated with the same adjective or noun.
The appropriate way to account for this non-independence is by using multi-level models and specifying so-called random effects [e.g., @gelman2006data; @schielzeth2009conclusions]. 
`r round(random_n[[3,1]],2) * 100`\% of linear models specified no random effects at all (without pooling their data), so effectively ignoring these non-independences [@hurlbert1984pseudoreplication]. <!-- TODO: We need to check if these model pooled -->
`r round(random_n[[1,1]],2) * 100`\% specified random intercepts only, and `r round(random_n[[2,1]],2) * 100`\% specified random intercepts and random slopes to account for the non-independence.
On average, teams that specified random effects, included `r round(rterms_n,1)` random terms in their models. 
Based on statistical framework, type of model, distribution family, fixed terms, and discarding random effects, there were a total of `r nrow(unique_model_specs)` different model specifications. 
<!-- TR: If done on team level did every team have a different mdoel?  -->

@@

Teams reviewed each others' acoustic and statistical analyses.
The mean rating of the acoustic analyses, on a scale from 0 to 100, is `r round(review_mean[[1]],1)` (SD = `r round(review_mean[[2]],1)`).
The mean rating of the statistical analysis is `r round(review_mean[[3]],1)` (SD = `r round(review_mean[[4]],1)`).
For reference, as mentioned in the Methods section, a score of 75 was defined as "an imperfect analysis but the needed changes are unlikely to dramatically alter the final interpretation", indicating that on average reviewers judged the provided analyses to be appropriate, although "imperfect".

### 2022-09-09 RR_manuscript/Draft_RR.tex
♻️ - edit descriptive results

In the following, we will describe the characteristics of our analyst teams and their models in order to give an overview over the sample and the large variation across analysis strategies.

\hypertarget{team-properties}{%
\subsubsection{Team properties}\label{team-properties}}

Eighty-Four teams initially signed up to participate.
Fifty-Four of the proposed teams dropped out of the project during the analysis phase\footnote{ADD REASONS}.
The remaining 30 teams submitted their analysis in time and were coded by the three initiating authors (SC, JC, TR).

Submitting teams consisted of an average of 3.63 analysts (sd = 2.8).
Analysts had on average 4.8 years (sd = 4.2) post PhD experience, ranging from -3.8 years, i.e.~PhD students (or younger) to 12.4 years.

Analysts prior belief in the effect under investigation ranged from 48.5 to 92 with an average of 69.2 (sd = 11.3), indicating that analysts had an overall rather high prior plausibility of the investigated relationship between prosody and typicality (scale ranged from 0 to 100).

The teams have submitted 170 models to answer the research question.
On average, teams submitted 3.6 models.
Crucially, analytic approaches widely varied in how they acoustically analysed the speech signal and how the statistically analysed the extracted values.

@@

Teams differed in what and how they measured the acoustic signal, including choosing different aspects of the acoustic signal, the temporal window over which they measured, and the concrete operationalization of how they measured those acoustic properties.
33\% of models used a duration measure as the outcome variable, 36\% used a f0 measure, 15\% used a formant measure, 14\% used an intensity measure, and 3\% used a different measure.

45\% of models measured acoustic properties at the level of the segment (e.g.~comparing the acoustic profile of a vowel), 44\% at the level of the word (e.g.~comparing the acoustic profile of ``banana''), 4\% at the level of the phrase (e.g.~the noun phrase including determiner and adjective, e.g.~``the green banana''), 2\% at the level of the whole sentence, and 5\% used a different time window. Based on a coarse coding of how acoustic measures were operationalized, we find a total of 55 different measurement specifications.

@@

The multiverse of acoustic measurement choices is exponentiated with choices during the statistical analysis, including the predictors and their operationalization, the chosen inferential framework, the type of model, and their model architecture, specifically how they accounted for dependencies in the data via random effect specifications in their multilevel models.

On average, models included 1.9 different predictors (counting only conceptually different predictors), i.e.~in addition to the critical predictor, typicality of the adjective noun combinations, most teams included additional predictors in their models. Predictors included the information structure of the sentence, trial number, semantic dimensions of the referent, part of speech, or gender of the speaker.

The original dataset allowed to operationalize the most relevant predictor, typicality, in different ways.
It was operationalized as categorical (e.g.~typical vs.~atypical) in 78\% of models, continuous (on a scale from 0-100) with the mean typicality rating in 19\% of models, and continuous with the median typicality rating in 1\% of models.

The majority of models were rooted in the frequentist framework (80\%). 20\% were operating in a Bayesian framework.
While teams almost exclusively used linear models to analyze their data (98\%, with a few exception using machine learning techniques or GAMs as a special case of a linear model), teams differed drastically in how they accounted for dependencies within the data.
The data contains several dependencies between data points, with multiple data points coming from the same subject and with multiple data points being associated with the same adjective or noun.
The appropriate way to account for this non-independence is by using multi-level models and specifying so-called random effects \citep[e.g.,][]{gelman2006data, schielzeth2009conclusions}.
7\% of linear models specified no random effects at all (without pooling their data), so effectively ignoring these non-independences \citep{hurlbert1984pseudoreplication}.
64\% specified random intercepts only, and 29\% specified random intercepts and random slopes to account for the non-independence.
On average, teams that specified random effects, included 2.4 random terms in their models.
Based on statistical framework, type of model, distribution family, fixed terms, and discarding random effects, there were a total of 45 different model specifications.

@@

Teams reviewed each others' analyses for both the acoustic analysis and the statistical analysis.
The mean rating of the quality of the acoustic analyses was 74.6 (sd = 11.1); the mean rating of the quality of the statistical analysis was 73.9 (sd = 14.1). For reference, the scale of the rating anchored a point grade of 75 as ``an imperfect analysis but the needed changes are unlikely to dramatically alter final interpretation'', indicating a rather high confidence of reviewers that the provided analyses yield appropriate (yet ``imperfect'') answers to the research question.

>>>>

In the following sections, we will describe the characteristics of the analysis teams that participated in the study and the analytic approaches they adopted.
An important aspect that emerges from the descriptive analysis is the large variation in analytic strategies.

\hypertarget{characteristics-of-analytic-teams}{%
\subsubsection{Characteristics of analytic teams}\label{characteristics-of-analytic-teams}}

Eighty-Four teams initially signed up to participate in the study.
Fifty-Four of the signed-up teams dropped out during the analysis phase\footnote{ADD REASONS}.
The remaining 30 teams submitted their analysis by the established deadline.
These are the teams that are included in this study.

Submitting teams consisted of an average of 3.63 individual analysts (SD = 2.8).

Upon sign-up, we collected background information from each analyst through the intake form, which was administered during Phase 1, prior to the distribution of the data to the teams.
Analysts had on average 4.8 years (SD = 4.2) of experience after completing their PhD, ranging from -3.8 years, i.e.~PhD students (or younger) to 12.4 years.
The analysts' prior belief in the effect under investigation on a scale from 0 to 100 ranged from 48.5 to 92 with an average of 69.2 (SD = 11.3).
We take this to suggest that analysts had an overall rather high positive prior belief in the investigated relationship between prosody and word combination typicality.

At the end of Phase 2 (primary data analysis), the teams have submitted a grand total 170 individual models to answer the research question, with a mean of 3.6 models per team.
Crucially, the analytic approaches employed by the teams varied widely both in the acoustical analyses of the speech signal and in the statistical analyses of the obtained measures.
We proceed to discuss these in the following sections, after which we provide an overview of the peer-review ratings.

@@

The analytic teams differed in their approach to the acoustic analysis of the speech signal, including choosing specific acoustic measures, the temporal window over which they are taken from and how the measures are transformed.
33\% of models used a measure of duration as the outcome variable, 36\% used fundamental frequency (f0), 15\% used vowel formants, 14\% intensity, and 3\% other measures.

45\% of models obtained acoustic measures at the level of the segment (e.g.~comparing the acoustic profile of a vowel), 44\% from the word level (e.g.~comparing the acoustic profile of \emph{Banane} `banana'), 4\% at the level of the phrase (e.g.~the noun phrase including determiner and adjective, e.g.~``the green banana''), 2\% from the whole sentence, and 5\% used a different time window.
Based on a coarse coding of how acoustic measures were operationalized, we find a total of 55 different measurement specifications.

@@

The high variability of acoustic measurements is multiplied by the choices of statistical analysis, including the chosen inferential framework, the type of model, and the model specification, like choice of predictors, interactions and group-level (aka random) effects.

The models included a mean of 2 different predictors (defined as variables or columns in the data table).
This means that, in addition to the critical predictor (typicality of the adjective noun combinations), there was on average at least an additional predictor.
Among the predictors included in the models there were information structure of the sentence, trial number, semantic dimensions of the referent, part of speech, or speaker gender.

The data given to the teams allowed them to operationalize the predictor of interest, word typicality, in different ways.
Among the possible operationalizations, 78\% of models contained typicality as a categorical variable (e.g.~atypical vs.~typical), 19\% used a continuous typicality scale from 0-100 by calculating the mean typicality for each word combination as obtained from the norming study, while 1\% of the models used the median typicality rating.

The majority of models were run within a frequentist framework (80\%).
20\% were run within a Bayesian framework.
While teams almost exclusively used linear models to analyze their data (98\%, with a few exception using machine learning techniques or GAMs as a special case of linear models), teams differed drastically in how they accounted for dependencies within the data.

The data contains several dependencies between data points, with multiple data points coming from the same subject and with multiple data points being associated with the same adjective or noun.
The appropriate way to account for this non-independence is by using multi-level models and specifying so-called random effects \citep[e.g.,][]{gelman2006data, schielzeth2009conclusions}.
7\% of linear models specified no random effects at all (without pooling their data), so effectively ignoring these non-independences \citep{hurlbert1984pseudoreplication}.
64\% specified random intercepts only, and 29\% specified random intercepts and random slopes to account for the non-independence.
On average, teams that specified random effects, included 2.4 random terms in their models.
Based on statistical framework, type of model, distribution family, fixed terms, and discarding random effects, there were a total of 45 different model specifications.

@@

Teams reviewed each others' acoustic and statistical analyses.
The mean rating of the acoustic analyses, on a scale from 0 to 100, is 74.6 (SD = 11.1).
The mean rating of the statistical analysis is 73.9 (SD = 14.1).
For reference, as mentioned in the Methods section, a score of 75 was defined as ``an imperfect analysis but the needed changes are unlikely to dramatically alter the final interpretation'', indicating that on average reviewers judged the provided analyses to be appropriate, although ``imperfect''.
